---
title: "Select an expanded list of compounds for the Turbyville experiment"
output: html_notebook
---

```{r message=FALSE}
library(magrittr)
library(platetools)
library(rvest)
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(progress = FALSE)
```

Plate map for the actual plate shipped to Dr. Turbyville is "C-7120-01-CMP-005.txt"
This file contains mapping of the Broad IDs in a 384 well plate that were sent.

Originally requested layout for that plate is "Broad Plate Map_final_modified_brds.csv".
I had to make some substitutions because few BRDs were not available in requested batch.


```{r message=FALSE}
first_batch_platemap <- read_tsv("../../docs/turbyville/C-7210-01-CMP-005.txt", col_names = T)
```

Plot the platemap

```{r}
library(platetools)
raw_map(data = ifelse(is.na(first_batch_platemap$broad_sample), "DMSO", "cpd"),
        well = first_batch_platemap$well_position,
        plate = 384) +
  ggtitle("C-7120-01-CMP-005") +
  theme_dark() +
  scale_fill_discrete()

```
Read list of compounds that were MTA'd

```{r message=FALSE}
mta <- read_csv("../../docs/turbyville/MTA2017412_broad_ids.txt")

mta_compounds <- 
  mta %>% 
  distinct(broad_sample) %>%
  mutate(pert_id = str_extract(broad_sample, "(BRD[-N][A-Z0-9]+)"))
```

Get list of compounds that were on the first Turbyville plate

```{r}

first_batch_compounds <-
  first_batch_platemap %>% 
  distinct(broad_sample) %>%
  mutate(pert_id = str_extract(broad_sample, "(BRD[-N][A-Z0-9]+)"))

```


Compounds that were MTA'd but not in the first plate

```{r}
mta_compounds %>% select(pert_id) %>%
  setdiff(first_batch_compounds %>% select(pert_id))
```

Compounds that were in the first plate but not MTA'd

```{r}
first_batch_compounds %>% select(pert_id) %>% 
  setdiff(mta_compounds %>% select(pert_id))
```

Create a full list of compounds - those that were MTA'd + those that were in the first plate

```{r}
mta_and_first_batch_compounds <-
  first_batch_compounds %>% select(pert_id) %>% 
  union(mta_compounds %>% select(pert_id))

```

Create a dictionary to get pert_id from compounds names

```{r}
Pf.cmpd <- readRDS("../results/master/2017-04-20_64917c8b/Pf_annotated.rds")
Pf.cmpd$data %<>% mutate(CPD_NAME = ifelse(CPD_NAME != "", CPD_NAME, Metadata_broad_sample))
cpdname_pertid <- 
  Pf.cmpd$data %>% 
  ungroup() %>%
  select(CPD_NAME, Metadata_broad_sample) %>%
  mutate(pert_id = str_extract(Metadata_broad_sample, "(BRD[-N][A-Z0-9]+)")) %>%
  select(-Metadata_broad_sample)
```



Get full list of compounds from  2017-05-03_b1f643c4 
https://broadinstitute.atlassian.net/wiki/spaces/IP/pages/114669056/2017-04-20+Final+list+of+gene+compound+connections
2017-04-20 Final list of gene compound connections
(for simple correlation/anti-correlation matches)

Search for 
KRAS_WT.1
KRAS_G12V
HRAS_G12V

```{r message=FALSE}

predictions <- 
  read_html("../results/HEAD/2018-08-05_b1f643c4/predictions.html") %>%
  html_nodes("table") %>%
  .[c(41,46,47)]

parse_tables <- function(index) {
  predictions[[index]] %>% 
    html_table(fill = TRUE) %>%
    dplyr::select(Metadata_broad_sample, 
           rep.corr,
           score = `Corr.`) %>%
    mutate(pert_id = str_extract(Metadata_broad_sample, "(BRD[-N][A-Z0-9]+)")) %>%
    dplyr::rename(broad_sample = Metadata_broad_sample) %>%
    inner_join(cpdname_pertid) %>%
    select(-CPD_NAME) %>%
    left_join(
      mta_compounds %>% select(pert_id) %>% mutate(is_mta = TRUE)
      ) %>% 
    mutate(is_mta = !is.na(is_mta))    
}

HRAS_G12V <- parse_tables(1)
KRAS_G12V <- parse_tables(2)
KRAS_WT <- parse_tables(3)

simple_correlations <-
  bind_rows(
    HRAS_G12V,
    KRAS_G12V,
    KRAS_WT
  )

```

How many unique compounds in this list?

```{r}
simple_correlations %>% distinct(pert_id) %>% count()
```

How many of these were MTA'd + first plate?

```{r}
simple_correlations %>%
  select(pert_id) %>%
  intersect(mta_and_first_batch_compounds) %>%
  count()
```

Verify that all the compounds listed in the for each of HRAS_G12V, KRAS_G12V, KRAS_WT (30 each) 
where in MTA + first plate
Result has 0 rows if true.

```{r}
simple_correlations %>%
  anti_join(mta_and_first_batch_compounds %>% select(pert_id))
```

Load list of compounds as listed in 
https://broadinstitute.atlassian.net/wiki/spaces/IP/pages/120356903/2017-08-24+Final+list+of+compounds+connected+to+HRAS+and+KRAS

```{r message=FALSE}
xCompound_List <- 
  read_csv("../../results/manual/2017-08-24_Final_list_of_compounds_connected_to HRAS_and_KRAS/Compound_List.csv") %>%
  select(broad_sample = Broad.ID) %>%
  arrange(broad_sample)

xKRAS_diff_MUT_WT <- 
  read_csv("../../results/manual/2017-08-24_Final_list_of_compounds_connected_to HRAS_and_KRAS/KRAS_diff_MUT_WT.csv") %>% 
  select(broad_sample = `Broad ID`) %>%
  arrange(broad_sample)

xKRAS_WT <- 
  read_csv("../../results/manual/2017-08-24_Final_list_of_compounds_connected_to HRAS_and_KRAS/KRAS_WT.csv") %>% 
  select(broad_sample = `Broad ID`) %>%
  arrange(broad_sample)

xKRAS_G12V <- 
  read_csv("../../results/manual/2017-08-24_Final_list_of_compounds_connected_to HRAS_and_KRAS/KRAS_G12V.csv") %>% 
  select(broad_sample = `Broad ID`) %>%
  arrange(broad_sample)

xHRAS_G12V <- 
  read_csv("../../results/manual/2017-08-24_Final_list_of_compounds_connected_to HRAS_and_KRAS/HRAS_G12V.csv") %>% 
  select(broad_sample = `Broad ID`) %>%
  arrange(broad_sample)
```

Is Compound_List the same as MTA'd compounds?

```{r}
all_equal(
  xCompound_List %>% select(broad_sample) %>% arrange(broad_sample),
  mta_compounds %>% select(broad_sample) %>%  arrange(broad_sample)
)
```

How many selected based on simple correlations?

```{r}
bind_rows(
   xKRAS_WT,
   xKRAS_G12V,
   xHRAS_G12V
   ) %>%
  distinct() %>%
  count()
```

How many more compounds are in the CSV files compared to those extracted from the html? (simple correlations)

```{r}
setdiff(
  bind_rows(
     xKRAS_WT,
     xKRAS_G12V,
     xHRAS_G12V
     ) %>%
    select(broad_sample) %>%
    distinct(),
  bind_rows(
     KRAS_WT,
     KRAS_G12V,
     HRAS_G12V
     ) %>%
    select(broad_sample) %>%
    distinct()  
) %>%
  count()

```
Check the same thing as above but on a per allele basis

KRAS_WT
```{r}
setdiff(
  xKRAS_WT %>%
    select(broad_sample) %>%
    distinct(),
  KRAS_WT %>%
    select(broad_sample) %>%
    distinct()
  ) %>%
  count()
```

KRAS_G12V
```{r}
setdiff(
  xKRAS_G12V %>%
    select(broad_sample) %>%
    distinct(),
  KRAS_G12V %>%
    select(broad_sample) %>%
    distinct()
  ) %>%
  count()
```

xHRAS_G12V
```{r}
setdiff(
  xHRAS_G12V %>%
    select(broad_sample) %>%
    distinct(),
  HRAS_G12V %>%
    select(broad_sample) %>%
    distinct()
  ) %>%
  count()
```

How many selected based on differential

```{r}
  distinct(
    xKRAS_diff_MUT_WT
    ) %>%
  count()
```

How many intersect?

```{r}
bind_rows(
   xKRAS_WT,
   xKRAS_G12V,
   xHRAS_G12V
   ) %>%
  distinct() %>%
  intersect(xKRAS_diff_MUT_WT) %>%
  distinct() %>%
  count()
```

How many total?

```{r}
bind_rows(
   xKRAS_WT,
   xKRAS_G12V,
   xHRAS_G12V,
   xKRAS_diff_MUT_WT) %>%
  distinct() %>%
  count()

distinct(xCompound_List) %>%
  count()
```

Get full list of compounds selected based on differential WT-mutant profiles

```{r message=FALSE}
differential_scores <- 
  read_csv("../../results/master/2018-08-08_a3c67c03/diff_cmpds.csv") %>%
  mutate(pert_id = str_extract(`compound ID`, "(BRD[-N][A-Z0-9]+)"))

```

Select top k differential profiles

```{r}
k <- 44 # 44 because that's what Mohammad set

differential_scores_top_k <- 
  differential_scores %>% 
  filter(abs(score) > 0.5) %>% 
  arrange(-score) %>%
  slice(1:k) %>% 
  select(pert_id) 

```

How many of these were MTA'd,  and MTA'd + first plate?

```{r}
differential_scores_top_k %>%
  select(pert_id) %>%
  intersect(mta_and_first_batch_compounds) %>%
  count()

differential_scores_top_k %>%
  select(pert_id) %>%
  intersect(mta_compounds %>% select(pert_id)) %>%
  count()
```

Do compounds selected based on 
- simple correlations, and
- differential scores
explain everything?

Result has 0 rows if true.

If result has non-zero rows, the compounds listed are the unaccounted ones

```{r}
mta_compounds %>%
  select(pert_id) %>%
  anti_join(simple_correlations %>% select(pert_id)) %>%
  anti_join(differential_scores_top_k)

```

Do the same as above, except using `simple_correlations_from_xls`

```{r}
simple_correlations_from_xls <-
  bind_rows(
     xKRAS_WT,
     xKRAS_G12V,
     xHRAS_G12V
     ) %>%
  select(broad_sample) %>%
  mutate(pert_id = str_extract(broad_sample, "(BRD[-N][A-Z0-9]+)")) %>%
  distinct()


mta_compounds %>%
  select(pert_id) %>%
  anti_join(simple_correlations_from_xls %>% select(pert_id)) %>%
  anti_join(differential_scores_top_k)

```

Plot the distribution of differential profile scores

```{r}

mta_differential_scores <- 
  differential_scores %>%
  rename(broad_sample = `compound ID`) %>%
  left_join(xKRAS_diff_MUT_WT %>% mutate(is_mta = TRUE)) %>%
  mutate(is_mta = !is.na(is_mta)) 

mta_differential_scores %>%
  ggplot(aes(is_mta, score)) + geom_jitter()

mta_differential_scores %>%
  filter(score > 0.5) %>%
  ggplot(aes(score, fill = is_mta)) + geom_histogram(binwidth = .01, alpha = 0.5)

```

Verify again that MTA'd compounds is exactly the same as the union of all the compounds in the xls files

```{r}

all_equal(
  bind_rows(
    xKRAS_WT,
    xKRAS_G12V,
    xHRAS_G12V,
    xKRAS_diff_MUT_WT) %>%
    distinct() %>%
    arrange(broad_sample),
  mta_compounds %>%
    select(broad_sample) %>%
    arrange(broad_sample)
  )
```


Looks like some compounds that were in the first batch had the same pert_id but different broad_id

```{r}

first_batch_compounds %>%
  distinct() %>%
  select(broad_sample) %>% 
  anti_join( 
    mta_compounds %>%
      select(broad_sample) %>%
      distinct()
    )

first_batch_compounds %>%
  distinct() %>%
  select(pert_id) %>% 
  anti_join( 
    mta_compounds %>%
      select(pert_id) %>%
      distinct()
    )
```


```{r}

differential_scores %<>% 
  select(broad_sample = `compound ID`,
         pert_id,
         score)
```


Create a new list of compounds based on differential profiles alone. We have a total of 95 inner wells and 76 edge wells to fill with these.

Create a list of compounds whose profiles are in the direction of mutant to WT. We want to pick 80% of (95 + 76) = 136 of these in total.

Some of these are already in the first batch

```{r}
differential_scores %>%
  filter(score > 0.5) %>% 
  inner_join(first_batch_compounds %>% select(pert_id)) %>%
  count()
```

We want to pick 80% of (95 + 76) = 136 of these in total, and 22 (result from the chunk above) have already been selected.
So we need to pick an additional 136 - 22 = 114
The bottom 76 / (95 + 76) = 44% = 60 of these should be placed on the edge. 

```{r}
differential_scores %>%
  filter(score > 0.5) %>% 
  anti_join(first_batch_compounds %>% select(pert_id)) %>%
  arrange(-score) %>%
  print() %>%
  write_csv("differential_compounds_going_from_mutant_to_wt_select_114_with_60_on_edge.csv")

```

Create a list of compounds whose profiles are in the direction of WT to mutant. We want to pick 20% of (95 + 76) = 34 of these.
None of these were in the first batch.

```{r}
differential_scores %>%
  filter(score < 0.5) %>% 
  inner_join(first_batch_compounds %>% select(pert_id)) %>%
  count()
```

The bottom 76 / (95 + 76) = 44% = 16 of these should be placed on the edge. 

```{r}
differential_scores %>%
  filter(score < 0.5) %>% 
  anti_join(mta_and_first_batch_compounds %>% select(pert_id)) %>%
  arrange(score) %>%
  print() %>%
  write_csv("differential_compounds_going_from_wt_to_mut_select_34_with_16_on_edge.csv")
```

Read available compounds

```{r}
available_compounds <- 
  read_csv("../../results/manual/2018-08-13_available_compounds/available_compounds.csv") %>%
  distinct(broad_sample) %>%
  mutate(pert_id = str_extract(broad_sample, "(BRD[-N][A-Z0-9]+)")) %>%
  mutate(is_available = TRUE)
  
differential_compounds_wt_to_mut <- 
  read_csv("../../results/master/2018-08-11_df8edeb6/differential_compounds_going_from_wt_to_mut_select_34_with_16_on_edge.csv") %>%
  left_join(available_compounds %>% select(pert_id, is_available)) %>%
  mutate(is_available = !is.na(is_available))

differential_compounds_mut_to_wt <- 
  read_csv("../../results/master/2018-08-11_df8edeb6/differential_compounds_going_from_mutant_to_wt_select_114_with_60_on_edge.csv") %>%
  left_join(available_compounds %>% select(pert_id, is_available)) %>%
  mutate(is_available = !is.na(is_available))

ggplot(differential_compounds_wt_to_mut, aes(-score)) + 
  geom_histogram(binwidth = 0.01) + 
  geom_rug(data = differential_compounds_wt_to_mut %>% filter(is_available),
           color = "red") + 
  ggtitle("WT to Mut", subtitle = "Red rug indicates available compounds.")

ggplot(differential_compounds_mut_to_wt, aes(score)) + 
  geom_histogram(binwidth = 0.01) + 
  geom_rug(data = differential_compounds_mut_to_wt %>% filter(is_available),
           color = "red") + 
  ggtitle("Mutant to WT", subtitle = "Red rug indicates available compounds.")

```


