---
title: "Inspect p38a screen"
output: html_notebook
---

```{r message=FALSE}
library(magrittr)
library(tidyverse)
```

Read data and drop `AdjP` because it was computed incorrectly.

```{r}
screen_data <- 
  read_csv("input/p38_data.csv") %>%
  select(-AdjP)
```
Recompute adjusted P-value from `KSP5`

```{r}
screen_data <-
  screen_data %>%
  mutate(AdjP = p.adjust(KSP5, method = "BH"))
```


- `is_rndm` indicates if the compounds is a Cell Painting negative control (based on correlation to the ORF)
- `medTLR5` is the ratio of the mean of the p38 imaging readout wrt DMSO for that compound (mean across the n=4 replicates)
- `Broad` is the Cell Painting correlation with the ORF 

```{r}
screen_data
```
Verify the semantics of `isrndm`

```{r}
screen_data %>%
  ggplot(aes(as.factor(isrndm), Broad)) + geom_point()
```
Create new variables

```{r}
screen_data <- 
  screen_data %>% 
  mutate(is_inhibitor = medTLR5 < 1)  %>%
  mutate(is_cp_negcon = as.factor(isrndm))
```

Plot KS-test adjusted p-value vs Cell Painting correlation

```{r}
p <-
  screen_data %>%
  ggplot(aes(Broad, -log10(AdjP), shape = is_cp_negcon, color = is_inhibitor)) +
  geom_point() +
  ggrepel::geom_text_repel(
    data = screen_data %>% filter(is_inhibitor),
    aes(label = sprintf("%0.2f", medTLR5)),
    max.overlaps = Inf,
    min.segment.length = unit(0, 'lines'),
    nudge_y = .3,
    color = "black",
    alpha = 0.5,
    size = 3
  ) +
  
  geom_hline(yintercept = -log10(0.05), linetype = "dotted") +
  coord_cartesian(clip = "off") +
  xlab("Cell Painting Correlation") +
  ylab("-log10 adjusted p-value") +
  theme_bw()

print(p)

ggsave("output/p38_figure_pvalue.png", width = 6, height = 4)
```

Plot ratio of medians of compounds:DMSO vs Cell Painting correlation

```{r}
p <- 
  screen_data %>%
  ggplot(aes(Broad, medTLR5, shape = is_cp_negcon, color = is_inhibitor)) + 
  geom_point() + 
  geom_text(
    aes(label = sprintf("%0.02f", medTLR5)), data = screen_data %>% filter(is_inhibitor)
  ) +
  coord_cartesian(clip = "off") +
  xlab("Cell Painting Correlation") +
  ylab("Ratio of medians") + 
  theme_bw()

print(p)

ggsave("output/p38_figure_ratio_median.png", width = 6, height = 4)
```

