---
title: "Inspect p38a single cell data"
output: html_notebook
---

```{r message=FALSE}
library(magrittr)
library(glue)
library(tidyverse)
```


```{r}
single_cell <-
  read_csv(
    "input/p38_single_cell.csv.gz",
    col_types = cols(
      P = col_integer(),
      R = col_integer(),
      C = col_integer(),
      Cmpd = col_character(),
      Conc = col_double(),
      Broad = col_double(),
      isrndm = col_logical(),
      KS5 = col_double(),
      KSP5 = col_double(),
      diff5 = col_double(),
      TLR5 = col_double()
    )
  )
```

- `is_rndm` indicates if the compounds is a Cell Painting negative control (based on correlation to the ORF)
- `Broad` is the Cell Painting correlation with the ORF 
- `KS5`: KS statistic 
- `KSP5`: KS p value
- `diff5`: ratio of median of compound to median of DMSO
- `TLR5`: raw p38 activity for that cell

```{r}
single_cell <-
  single_cell %>%
  mutate(is_dmso = Cmpd == "DMSO")
```

No perturbations in P=10

```{r}
single_cell %>%
  filter(!is_dmso) %>%
  count(P)
```
```{r}
single_cell <-
  single_cell %>%
  filter(P != 10)
```

~99.9% of the values are < 1.05

```{r}
single_cell %>%
  group_by(P) %>%
  summarise(q999 = quantile(TLR5, c(.999))) %>%
  arrange(desc(q999))
```

```{r}
single_cell <-
  single_cell %>%
  filter(TLR5 < quantile(TLR5, 0.999))
```

>  p38 activity in single cells was calculated using the ratio of the median intensity of the p38-KTR in a 5-pixel-wide cytoplasmic ring around the nucleus to the median intensity of the p38-KTR in the nucleus.

```{r}
single_cell <- 
  single_cell %>%
  filter(TLR5 > 0) 
```


```{r}
single_cell %>%
  ggplot(aes(TLR5, fill = is_dmso)) + 
  geom_histogram(position="identity", alpha = 0.5, bins = 50) + 
  facet_wrap(~P, ncol = 3, scales = "free_x")
```

There are two DMSO's in each of the 10 columns

```{r}
single_cell %>%
  filter(is_dmso) %>%
  distinct(P, R, C) %>%
  count(P, C) %>%
  ggplot(aes(C, n)) + 
  geom_point() + 
  facet_wrap(~P) +
  xlab("Column") +
  ylab("Number of DMSO wells")
```

There's some evidence of column effect

```{r}
single_cell %>%
  filter(is_dmso) %>%
  ggplot(aes(TLR5, group = C)) + geom_density() + facet_wrap(~P)
```


```{r}
single_cell %>%
  filter(is_dmso) %>%
  group_by(P, C) %>%
  summarise(mean_TLR5 = mean(TLR5)) %>%
  ggplot(aes(as.factor(P), mean_TLR5)) + 
  geom_boxplot() +
  ggtitle("Distribution of mean TLR5 values in DMSO wells")
```

Likely not much of row effect (but hard to know because it's just n = 2)

```{r}
single_cell %>%
  filter(is_dmso) %>%
  ggplot(aes(TLR5, group = R)) + geom_density() + facet_wrap(~P)
```

Normalizing w.r.t. DMSO in the same column is therefore justified.

Inspect the plate layout.

```{r}
single_cell %>%
  distinct(P, R, C, Cmpd) %>%
  ggplot(aes(C, R, fill = Cmpd)) + 
  geom_tile(color = "black") + 
  coord_equal() + 
  theme_minimal() + 
  facet_wrap(~P) +
  theme(legend.position = "none")
```
A compound is present only on a single plate

```{r}
single_cell %>%
  filter(!is_dmso) %>%
  distinct(P, Cmpd) %>%
  count(Cmpd) %>% 
  filter(n > 1)
```

Trim DMSOs

```{r}
single_cell <-
  single_cell %>%
  inner_join(single_cell %>%
               filter(!is_dmso) %>%
               distinct(P, C))
```

Inspect the plate layout.

```{r}
single_cell %>%
  distinct(P, R, C, Cmpd) %>%
  ggplot(aes(C, R, fill = Cmpd)) + 
  geom_tile(color = "black") + 
  coord_equal() + 
  theme_minimal() + 
  facet_wrap(~P) +
  theme(legend.position = "none")
```


```{r}
compute_statistics <- function(population, sample, variable) {
  
  population %>%
    group_by(P, R, C, Cmpd, Broad, isrndm) %>%
    summarise(ks_pval = ks.test(.data[[variable]], sample[[variable]])$p.value,
              ks_stat = ks.test(.data[[variable]], sample[[variable]])$statistic,
              t_pval = t.test(.data[[variable]], sample[[variable]])$p.value,
              t_stat = t.test(.data[[variable]], sample[[variable]])$statistic,
              median_ratio = mean(.data[[variable]]) / mean(sample[[variable]]), 
              .groups = "keep")
}
```


```{r}
statistics <-
  cytominer::stratify(
  population = single_cell %>% filter(!is_dmso),
  sample = single_cell %>% filter(is_dmso),
  reducer = compute_statistics,
  strata = c("P", "C"),
  variable = "TLR5"
)
```

Statistics are computed per replicate (n=4) for each of the 34 compounds

```{r}
statistics %>%
  count(Cmpd) %>%
  count(n)
```

```{r}
statistics %>%
  ggplot(aes(Cmpd, t_stat)) + 
  geom_boxplot() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + 
  facet_wrap(~isrndm, scales = "free_x")
```
```{r}
statistics %>%
  ggplot(aes(Cmpd, median_ratio)) + 
  geom_boxplot() +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + 
  facet_wrap(~isrndm, scales = "free_x")
```
```{r eval=TRUE}
plot_distribution <- function(Cmpdi) {
  statistics_i <-
    statistics %>%
    filter(Cmpd == Cmpdi) %>%
    summarise(across(all_of(
      c("ks_pval", "ks_stat", "t_pval", "t_stat", "Broad", "isrndm")
    ), mean))
  
  cp_type <- ifelse(statistics_i$isrndm == 1, "cpneg", "cphit")
  cmpd_single_cell <-
    single_cell %>%
    ungroup() %>%
    filter(Cmpd == Cmpdi)
  
  dmso_single_cell <-
    cmpd_single_cell %>%
    distinct(P, C) %>%
    inner_join(single_cell %>% filter(is_dmso), by = c("P", "C"))
  
  combined_single_cell <-
    bind_rows(dmso_single_cell,
              cmpd_single_cell)
  
  p <-
    combined_single_cell %>%
    ggplot(aes(TLR5, fill = Cmpd)) +
    geom_density(position = "identity",
                   alpha = 0.5) +
    facet_wrap( ~ C) +
    ggtitle(glue("{cp_type}_{Cmpdi}"),
            subtitle =
              with(
                statistics_i,
                sprintf(
                  "ks_stat = %.2f ks_pval = %.2g t_stat = %.1f t_pval = %.2g cp_cor = %.2f",
                  ks_stat,
                  ks_pval,
                  t_stat,
                  t_pval,
                  Broad
                )
              )) 
  
  ggsave(
    glue("output/distributions/{cp_type}_{Cmpdi}.png"),
    p,
    width = 6,
    height = 4
  )
}
```


```{r eval=TRUE}
statistics %>%
  distinct(Cmpd) %>%
  pull("Cmpd") %>%
  walk(plot_distribution)
```


```{r}
statistics_agg <-
  statistics %>%
  group_by(Cmpd) %>%
  summarise(across(all_of(
    c("ks_pval", "ks_stat", "t_pval", "t_stat", "median_ratio", "Broad", "isrndm")
  ), mean)) %>%
  mutate(t_pval_adj = p.adjust(t_pval, "BH"),
         ks_pval_adj = p.adjust(ks_pval, "BH"))
```


```{r}
statistics_agg %>%
  ggplot(aes(Broad, -log10(t_pval_adj))) + 
  geom_point() + 
  geom_hline(yintercept = -log10(0.01)) + 
  xlab("Cell Painting Correlation") +
  coord_cartesian(clip = "off")
```
```{r}
statistics_agg %>% 
  mutate(is_significant_t_test_01 = t_pval_adj < 0.01) %>%
  mutate(is_cp_negative_control = isrndm == 1) %>%
  count(is_cp_negative_control, is_significant_t_test_01)
```


```{r}
statistics_agg %>%
  ggplot(aes(Broad, -log10(ks_pval_adj))) + 
  geom_point() + 
  geom_hline(yintercept = -log10(0.01)) + 
  xlab("Cell Painting Correlation") +
  coord_cartesian(clip = "off")
```
```{r}
statistics_agg %>% 
  mutate(is_significant_ks_test_01 = ks_pval_adj < 0.01) %>%
  mutate(is_cp_negative_control = isrndm == 1) %>%
  count(is_cp_negative_control, is_significant_ks_test_01)
```

```{r}
statistics_agg %>%
  ggplot(aes(Broad, median_ratio)) + 
  geom_point() + 
  xlab("Cell Painting Correlation") +
  coord_cartesian(clip = "off")
```


```{r}
statistics_agg %>%
  ggplot(aes(Broad, t_stat)) + 
  geom_point() +
  xlab("Cell Painting Correlation") +
  coord_cartesian(clip = "off")
```


```{r}
statistics_agg %>%
  ggplot(aes(Broad, ks_stat)) + 
  geom_point() +
  xlab("Cell Painting Correlation") +
  coord_cartesian(clip = "off")
```
```{r}
screen_data <- 
  read_csv("input/p38_data.csv") 
```
```{r}
inner_join(statistics_agg %>% select(Cmpd, ks_pval),
           screen_data %>% select(Cmpd, KSP5)) %>%
  ggplot(aes(ks_pval, KSP5)) +
  geom_point() + 
  coord_equal() +
  geom_abline(slope = 1) +
  xlab("ks p-value (R)") +
  xlab("ks p-value (Matlab)")
```

