---
title: "Inspect PGC1a screen"
output: html_notebook
---

```{r message=FALSE}
library(magrittr)
library(tidyverse)
```


Trail that led me to the data

1. [Paper](https://www.cell.com/cell/fulltext/S0092-8674(17)30249-0#secsectitle0075)
2. [Project page](https://pubchem.ncbi.nlm.nih.gov/bioassay/651721)
3. [Dataset within the project](https://pubchem.ncbi.nlm.nih.gov/bioassay/651723#section=Data-Table)

The "Download data table" and you get this URL below

```{r}
screen <- read_csv("https://pubchem.ncbi.nlm.nih.gov/assay/pcget.cgi?query=download&record_type=datatable&actvty=all&response_type=save&aid=651723")
```

```{r}
set.seed(42)
screen %>% 
  sample_n(10)
```


```{r}
screen %>%
  ggplot(aes(PUBCHEM_ACTIVITY_SCORE, fill = PUBCHEM_ACTIVITY_OUTCOME)) + 
  geom_histogram(bins = 100) + 
  theme(legend.position = "bottom")

screen %>%
  ggplot(aes(PUBCHEM_ACTIVITY_OUTCOME, PUBCHEM_ACTIVITY_SCORE, )) + 
  geom_boxplot()


screen %>%
  count(PUBCHEM_ACTIVITY_OUTCOME)
```

Save the CIDs and then query https://pubchem.ncbi.nlm.nih.gov/pc_fetch/pc_fetch.cgi

```{r eval=FALSE}
screen %>%
  select(PUBCHEM_CID) %>%
  na.omit() %>%
  write_tsv("~/Downloads/pgc1a_screen_cid.txt", col_names = FALSE)
```

The results were saved in `input/3843044851509932684.txt.gz`

```{r}
smiles_cid <- read_tsv("input/3843044851509932684.txt.gz", col_names = c("PUBCHEM_CID", "CBP_SMILES"))
```
```{r}
screen %>%
  inner_join(smiles_cid) %>%
  write_csv("input/pgc1a_screen.csv.gz")
```
See this
https://github.com/broadinstitute/2016_08_22_Gene_Compound/issues/6#issuecomment-870740194

```{r}
screen_internal <- read_csv("input/puma_assay_data_CBIP_2139-01.csv.gz")
```
Compare the distributions

```{r}
screen %>% ggplot(aes(PUBCHEM_ACTIVITY_SCORE)) + geom_histogram(bins = 100)
screen_internal %>% ggplot(aes(RESULT_VALUE)) + geom_histogram(bins = 100)
```
The PUMA version of the readouts are  normalized from -1 to 1

> We prepared assay performance profiles following a double sigmoid normalization procedure to ensure that all readouts are scaled in the same range
> https://www.biorxiv.org/content/10.1101/2020.12.15.422887v1.full
